\documentclass[a4paper, twoside]{report}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{hyperref}
\hypersetup{colorlinks=false}
\usepackage{lscape}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{setspace}
\usepackage{subcaption}


\onehalfspacing
%% Sets page size and margins
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\title{Metamaterials, Non-affine Deformations and Deployable Structures}
\author{Apoorv Srivastava}
% Update supervisor and other title stuff in title/title.tex

\begin{document}
\input{title/title.tex}

\begin{abstract}
Deployable Structures are the structures which can be collapsed into compact modules and deployed when needed. While unlike their static counterparts, these structures are not ubiquitous, some examples of deployable structures present around us are Umbrellas, Scissors and foldable furniture such as tables and chairs. Affine deformations are deformations are the deformations which translate uniformly at the microscopic level and can be defined using linear translations and rotations, following this, non-affine deformation are the deformations which do not translate linearly from microscopic to macroscopic level owing to the topology of the materials. Coupling the ideas or non-affine deformations and deployable structure at microscopic level leads to design of materials having special properties and such materials are called metamaterials. In addition to these, the project involved study of Mechanism, Lattices, Origami and Topological Mechanics.\vspace{8 pt}


The later half of the project involved modelling the plates using the spring ball models and benchmarking the results with analytical theories. This was done in order to develop a program which can accurately model plates, following which study of the effect of pre-stress in the springs on the properties of the plate can be carried out. This will help in determining strain fields which will result in plate deforming to a specified shape and the ways in which pre-stress affect the dynamic properties of plate. An added advantage of such modelling is that it can be used to predict the plate displacements and following that stresses in the plate for complex loading and cases where theoretical results are not available or are difficult to arrive at.


\end{abstract}

\renewcommand{\abstractname}{Acknowledgments}
\begin{abstract}
I would like to thank Prof. Mandar. M. Inamdar for connecting me to all these fascinating topics and insightful discussions which proved to be of great help in understanding these topics. His advices have been instrumental in this project and have helped me immensely in tackling the most obscure problems I had come across. I find myself very fortunate to find his guidance. I would also like to thank Prof. Venkata S. K. Delhi for connecting me to the exciting area of research and Kshitij Vijaywargiya for initially uncovering the idea of deployable structures for me. \vspace{8 pt}

Special thanks to my wing mates for being a constant source of support and entertainment, everytime I was stressed. I am deeply indebted to my family for their constant and unconditional support in this endeavour.
\end{abstract}

\tableofcontents
\listoffigures

\input{introduction/introduction.tex}
\input{Work/Origami_Model.tex}
\input{Work/Plate_Model.tex}
\input{Work/Plate_Model Thickness study and Strain field}
\input{FutureWork/FutureWork.tex}
\input{appendix/appendix.tex}

\bibliographystyle{unsrt}
\bibliography{bibs/sample}
\addcontentsline{toc}{chapter}{Bibliography}

\end{document}